This is the title [Article v\versionnumber]

### 2022-08-19 manuscript/manuscript.tex

\subsection{Software/system requirements}
%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
The main implementation of GIST is part of the MD analysis software \software{cpptraj} as part of AmberTools \cite{amber22}.
A recent version of \software{cpptraj} should be used, such as the one in AmberTools22 or an installation from GitHub.
The closed-source Amber simulation engine is used in this tutorial to run the necessary MD simulations.

If no Amber installation is available, the MD can be performed other simulation engines compatible with \software{cpptraj}.
If using GROMACS, we propose to prepare the structures and topologies with AmberTools and converting them using \software{acpype.py}.
In that way, a full Amber topology is available for GIST.
Also note that compressed trajectory formats such as \inlinecode{xtc} might bias the entropy calculation.

Furthermore, a recent Python version (>3.6) should be available with the following packages.
\begin{itemize}
	\item \software{gisttools} (and its prerequisites)
	\item \software{mdtraj}
\end{itemize}

@@

If the reader prefers to skip the MD and GIST calculation, we also provide the GIST output files with this tutorial, such that the post-processing can be done without any expensive calculations. We also provide a Jupyter notebook \cite{Kluyver2016-jupyter,Granger2021-jupyter} with the presented analyses.

\section{Theory}
GIST is an implementation of Inhomogeneous Solvation Theory (IST) \cite{Lazaridis1998} that discretizes  the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} to overcome the limitation of IST which was commonly used for regions of high water occupancy and its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST. \cite{Kraml2020,Chen2021}

@@

\begin{equation}
	\Delta S_\textit{solv}
	\approx \Delta S_\textit{sw}
	\newcommand{\gro}{g_{sw}(\textbf{r},\omega)}
	\equiv -k_\textit{B} \frac{\rho^\textit{0}}{8\pi^\textit{2}} \int \gro \ln\gro d\textbf{r}d\omega
\end{equation}
\todo{We are inconsistent regarding R/kB}

@@

The first method assumes that the position ($r$) is independent of the orientation ($\omega$). This approximation allows for the 
splitting of the contributions of the translational and the orientational entropy. The nearest neighbor (NN) approach is used to 
evaluate each expression where $N_\textit{k}$ is the number of water molecules found in voxel k, $\gamma$ is the Euler's constant 
that accounts for the bias in naive entropy estimator, and $g_\textit{NN}$ is the nearest neighbor estimate. In the nearest neighbor
estimate, a voxel is considered to be a neighboring voxel when it shares any vertices of the voxel of interest. 
\begin{equation}
	g_\textit{vox} \left( \textbf{r}, \omega \right) \approx g_\textit{vox} \left( \textbf{r})g_\textit{vox}\omega \right)
\end{equation}

\begin{equation}
	S_{k}^\textit{trans} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} ln g_{NN, \textit{i}} \left( \textbf{r} \right) + \gamma \right)
\end{equation}

@@

The second method directly calculates the solvation entropy by evaluating the six-dimensional integral (3 for position and 3 for orientation) 
using the nearest-neighbor approach. Unlike the first method, the nearest neighbor estimate for the sixth-order entropy considers all the water 
found within the same and neighboring voxels.  
\begin{equation}
	S_\textit{k} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} ln g_{NN, \textit{i}} \left( \textbf{r}, \omega \right) + \gamma \right)
\end{equation}

@@

\begin{equation}
	E_\textit{lj} = E_\textit{lj\_short} +  E_\textit{corr}
\end{equation}

@@

Here, we choose not to include any sampling of the conformational states. 
It is therefore important that the conformations of the monomers in the complex and dissociated states match exactly. 
Use \software{cpptraj} to produce biotin and streptavidin structures based on the complex by stripping the respective other molecule like this:
\begin{lstlisting}
parm complex/solvated.parm7
trajin complex/equil/EQUIL-DONE.rst
strip ^2 parmout streptavidin/solvated.parm7
trajout streptavidin/pre-equil.ncrst
\end{lstlisting}
(Adapt for biotin as necessary.)

%We provide a script called \software{cpptraj\_remove\_mol.sh} that removes a single molecule from a topology and structure using \software{cpptraj} without altering the rest of the system. \todo{Explain cpptraj rather than our script. The script is for the automated workflow only.}

@@

The expected output of this code is shown in Figure~\ref{fig_radial_convergence}.
Both the solvent-solvent energy and the entropy are not perfectly flat, indicating that the reference value is not optimal.
The slow energy convergence indicates that a non-optimal reference value was used, but is also complicated by the fact that a charged solute is used.
In this case, both the entropy and $E_{sw}$ can be considered converged around \SI{10}{\angstrom}.
While this is not true for $E_{ww}$, we will see below that the energy difference upon binding converges more readily than its individual contributions.
%\todo{The reader is left hanging here - what should be done about the convergence issue? Is that bad or can it be ignored? -vah} - ok? -fwa

>>>>

\subsection{Software/system requirements}
%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
The main implementation of GIST is part of the MD analysis software \software{cpptraj} as part of AmberTools \cite{amber22}.
A recent version of \software{cpptraj} should be used, such as the one in AmberTools22 or an installation from GitHub (\url{https://github.com/Amber-MD/cpptraj}).
The closed-source Amber simulation engine is used in this tutorial to run the necessary MD simulations.

If no Amber installation is available, the MD can be performed other simulation engines compatible with \software{cpptraj}.
If using GROMACS, we propose to prepare the structures and topologies with AmberTools and converting them using \software{acpype.py}.
In that way, a full Amber topology is available for GIST. 
\todo{ (vah) is it necessary to mention GROMACS specifically? I would remove the previous two lines}
Also note that compressed trajectory formats such as \inlinecode{xtc} might bias the entropy calculation.

Furthermore, a recent Python version (>3.6) should be available with the following packages.
\begin{itemize}
	\item \software{gisttools} (\url{https://github.com/liedllab/gisttools})
	\item \software{mdtraj (\url{https://www.mdtraj.org/})}
\end{itemize}

@@

If the reader prefers to skip the MD and GIST calculation, we also provide the GIST output files with this tutorial, such that the post-processing can be done without any expensive calculations.
We also provide a Jupyter notebook \cite{Kluyver2016-jupyter,Granger2021-jupyter} with the presented analyses. 
We recommend JupyterLab to work with the notebook. Jupyter is available at \url{https://jupyter.org/}.

For the visualization of three-dimensional contributions we require molecular visualization software that can handle \software{OpenDX} files.
We recommend using a current version of \software{PyMol}\cite{pymol} or \software{VMD}\cite{vmd}.
\section{Theory}
GIST is an implementation of Inhomogeneous Solvation Theory (IST) \cite{Lazaridis1998} that discretizes  the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} to overcome the limitation of IST which was commonly used for regions of high water occupancy. 
Its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST. \cite{Kraml2020,Chen2021}

@@

\begin{equation}
	\Delta S_\textit{solv}
	\approx \Delta S_\textit{sw}
	\newcommand{\gro}{g_{sw}(\textbf{r},\omega)}
	\equiv -R \frac{\rho^\textit{0}}{8\pi^\textit{2}} \int \gro \ln\gro d\textbf{r}d\omega
\end{equation}
\todo{We are inconsistent regarding R/kB; vah: Fixed. is it - tho? Papers seem to disagre }

@@

The first method assumes that the position ($r$) is independent of the orientation ($\omega$). This approximation allows for the 
splitting of the contributions of the translational and the orientational entropy. The nearest neighbor (NN) approach is used to 
evaluate each expression where $N_\textit{k}$ is the number of water molecules found in voxel k, $\gamma$ is the Euler's constant 
that accounts for the bias in naive entropy estimator, and $g_\textit{NN}$ is the nearest neighbor estimate. In the nearest neighbor
estimate, a voxel is considered to be a neighboring voxel when it shares any vertices of the voxel of interest. 
\todo{ vah: Is gamma divided by Nk or not? The formula in Franz' recent paper indicates otherwise.}
\begin{equation}
	g_\textit{vox} \left( \textbf{r}, \omega \right) \approx g_\textit{vox} \left( \textbf{r})g_\textit{vox}\omega \right)
\end{equation}

\begin{equation}
	S_{k}^\textit{trans} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}} \left( \textbf{r} \right) + \gamma \right)
\end{equation}

@@

The second method directly calculates the solvation entropy by evaluating the six-dimensional integral (3 for position and 3 for orientation) 
using the nearest-neighbor approach. Unlike the first method, the nearest neighbor estimate for the sixth-order entropy considers all the water 
found within the same and neighboring voxels.  
\begin{equation}
	S_\textit{k} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}} \left( \textbf{r}, \omega \right) + \gamma \right)
\end{equation}

@@

\begin{equation}
	E_\textit{lj} = E_\textit{lj,\ short} +  E_\textit{corr}
\end{equation}

@@

Here, we choose not to include any sampling of the conformational states. 
It is therefore important that the conformations of the monomers in the complex and dissociated states match exactly. 
Use \software{cpptraj} to produce biotin and streptavidin structures based on the complex by stripping the respective other molecule like this:
\begin{lstlisting}
parm complex/solvated.parm7
trajin complex/equil/EQUIL-DONE.rst
strip ^2 parmout streptavidin/solvated.parm7
trajout streptavidin/pre-equil.ncrst
\end{lstlisting}
(Adapt for biotin as necessary.)
\todo{vah: Maybe include a section on how to prepare a protein structure before equilibration. Protonation, force field and a large enough iso solvation box seem important to me. I will probably also at least mention what simulation settings we are running later on.}
%We provide a script called \software{cpptraj\_remove\_mol.sh} that removes a single molecule from a topology and structure using \software{cpptraj} without altering the rest of the system. \todo{Explain cpptraj rather than our script. The script is for the automated workflow only.}

@@

The expected output of this code is shown in Figure~\ref{fig_radial_convergence}.
Both the solvent-solvent energy and the entropy are not perfectly flat, indicating that the reference value is not optimal.
The slow energy convergence indicates that a non-optimal reference value was used, but is also complicated by the fact that a charged solute is used.
In this case, both the entropy and $E_{sw}$ can be considered converged around \SI{10}{\angstrom}.
While this is not true for $E_{ww}$, we will see below that the energy difference upon binding converges more readily than its individual contributions.
\todo{vah: Is that shown somewhere yet?}
